### 2022-09-14 Resources/LogJournal.Rmd
first commit

---
title: "Log"
author: "J.A. Busker"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
```

<!-- Copyright (c) 2022 John Busker. -->
<!-- Licensed under GPLv3. See gpl.md -->

---

The data has been retrived from [uci.](https://archive.ics.uci.edu/ml/datasets/Yeast)

Attributes information:

1.  Sequence Name: Accession number for the SWISS-PROT database  
2.  mcg: McGeoch's method for signal sequence recognition.  
3.  gvh: von Heijne's method for signal sequence recognition.  
4.  alm: Score of the ALOM membrane spanning region prediction program.  
5.  mit: Score of discriminant analysis of the amino acid content of the N-terminal region (20 residues long) of mitochondrial and non-mitochondrial proteins.  
6.  erl: Presence of "HDEL" substring (thought to act as a signal for retention in the endoplasmic reticulum lumen). Binary attribute.  
7.  pox: Peroxisomal targeting signal in the C-terminus.  
8.  vac: Score of discriminant analysis of the amino acid content of vacuolar and extracellular proteins.  
9.  nuc: Score of discriminant analysis of nuclear localization signals of nuclear and non-nuclear proteins.  
10. Class Distribution: The class is the localization site. Consisting of:

        CYT (cytosolic or cytoskeletal)                    463  
        NUC (nuclear)                                      429  
        MIT (mitochondrial)                                244  
        ME3 (membrane protein, no N-terminal signal)       163  
        ME2 (membrane protein, uncleaved signal)            51  
        ME1 (membrane protein, cleaved signal)              44  
        EXC (extracellular)                                 37  
        VAC (vacuolar)                                      30  
        POX (peroxisomal)                                   20  
        ERL (endoplasmic reticulum lumen)                    5

```{r}
data <- as_tibble(read.table("yeast.data", sep = ""))
```

### 2022-09-16 Resources/LogJournal.Rmd
Added the column names to the dataset
Made my own codebook and loaded in the data

@@
 
-1.  Sequence Name: Accession number for the SWISS-PROT database  
+1.  Sequence Name: Accession number for the SWISS-PROT database.  
 2.  mcg: McGeoch's method for signal sequence recognition.  
@@
 9.  nuc: Score of discriminant analysis of nuclear localization signals of nuclear and non-nuclear proteins.  
-10. Class Distribution: The class is the localization site. Consisting of:
+10. Class Distribution: The class is the localization site. Consisting of (abbreviation (full name) the amount):
 
@@
 
+## EDA of the yeast data
+
+Since there is not a codebook. I created my own codebook.
+
+```{r}
+attr_names <- c("seq.name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac","nuc", "loc.site" )
+data_types <- c("str", "int", "int", "int", "int", "int", "bool", "int", "int", "factor")
+data_labels <- c("Accession number for the SWISS-PROT database",
+                                    "McGeoch's method for signal sequence recognition", 
+                                    "von Heijne's method for signal sequence recognition", 
+                                    "Score of the ALOM membrane spanning region prediction program",
+                                    "Score of discriminant analysis of the amino acid content of the N-terminal region",
+                                    "Presence of 'HDEL' substring",
+                                    "Peroxisomal targeting signal in the C-terminus",
+                                    "Score of discriminant analysis of the amino acid content of vacuolar and extracellular proteins",
+                                    "Score of discriminant analysis of nuclear localization signals of nuclear and non-nuclear proteins",
+                                    "The class is the localization site")
+codebook <- data.frame(Name=attr_names, Fullname=data_labels, Datatypes=data_types)
+
+knitr::kable(codebook)
+```
+
+Here is the codebook. With the attributes abbreviation, explanation and data types. As you can see there are many int types.
+
+Now we need to load in the data. And change the column names to the attributes abbreviation, since these are non-existent.
+
 ```{r}
 data <- as_tibble(read.table("yeast.data", sep = ""))
+colnames(data) <- attr_names
+
+str(data)
 ```
+
+As you can see, the last column, loc.site, consists of characters, but this one needs to be converted to factors for clarity and complexity.
+
+```{r}
+data$loc.site <- as.factor(data$loc.site)
+str(data$loc.site)
+```
+
+As you can see the data type of the column loc.site has been succesfully changed to factors.

### 2022-09-20 Resources/LogJournal.Rmd
Added more info about the dataset
Added a violion plot with jitter points

--- 2022-09-14 Resources/LogJournal.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, comment=NA)
 library(tidyr)
+library(kableExtra)
+library(ggplot2)
+library(gridExtra)
+library(cowplot)
 ```
@@
 
-The data has been retrived from [uci.](https://archive.ics.uci.edu/ml/datasets/Yeast)
+The data has been retrived from [uci.](https://archive.ics.uci.edu/ml/datasets/Yeast) The dataset has been gathered of 1484 yeast sequences from SWISS-PROT using the annotations from YPD.
 
@@
 
+## EDA of the yeast data
+
+Since there is not a codebook. I created my own codebook.
+
+```{r}
+attr_names <- c("seq.name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac","nuc", "loc.site" )
+data_types <- c("str", "float", "float", "float", "float", "float", "bool", "float", "float", "factor")
+data_labels <- c("Accession number for the SWISS-PROT database",
+                                    "McGeoch's method for signal sequence recognition", 
+                                    "von Heijne's method for signal sequence recognition", 
+                                    "Score of the ALOM membrane spanning region prediction program",
+                                    "Score of discriminant analysis of the amino acid content of the N-terminal region",
+                                    "Presence of 'HDEL' substring",
+                                    "Peroxisomal targeting signal in the C-terminus",
+                                    "Score of discriminant analysis of the amino acid content of vacuolar and extracellular proteins",
+                                    "Score of discriminant analysis of nuclear localization signals of nuclear and non-nuclear proteins",
+                                    "The class is the localization site")
+codebook <- data.frame(Name=attr_names, Fullname=data_labels, Datatypes=data_types)
+
+codebook %>%
+  kbl(caption = "Codebook of the data") %>%
+  kable_classic(html_font = "Cambria")
+```
+
+Here is the codebook. With the attributes abbreviation, explanation and data types. As you can see there are many float datatypes.
+
+Now we need to load in the data. And change the column names to the attributes abbreviation, since these are non-existent. Let's give all the columns a name. And let's take a quick look at the data.
+
 ```{r}
 data <- as_tibble(read.table("yeast.data", sep = ""))
+colnames(data) <- attr_names
+head(data)
 ```
+
+We can drop the first columns since it is not necessary.
+
+```{r}
+data <- data[, -1]
+head(data)
+```
+
+The first column has been succesfully dropped from the data.
+
+We are going to look for any missing values.
+
+```{r}
+sum(is.na(data))
+```
+
+There are not any missing values.
+
+Now we need to take a clearer look at the data set using `str()`
+
+```{r}
+str(data)
+```
+
+As you can see, the last column, loc.site, consists of characters, but this one needs to be converted to factors for clarity and complexity.
+
+```{r}
+data$loc.site <- as.factor(data$loc.site)
+str(data$loc.site)
+```
+
+As you can see the data type of the column loc.site has been succesfully changed to factors.
+
+The ERL variable also should be changed. As you can see, it is now a num data type but I needs to be a bool/binary datatype. Let's take a closer look at th ERL column.
+
+```{r}
+summary(data$erl)
+```
+
+This looks weird because the value should be 0 or 1.
+
+There are alot of 0.5 values. But should not exist in this column. The variable need to be a bool, or in other words it must be a 0 or a 1. So all the 0.5 values needs to be changed to a 0.
+
+```{r}
+table(data[, "erl"])
+```
+
+Before the data transformation there are 1470 counts of the value 0.5 and 14 of 1.
+
+```{r}
+data$erl[data$erl == 0.5] <- 0
+table(data[,"erl"])
+```
+
+The data has been succesfully transformed. Now the datatype has to be changed to a bool.
+
+```{r}
+data$erl <- as.logical(data$erl)
+str(data$erl)
+```
+
+Now every column has the right datatype.
+
+```{r}
+p1 <- ggplot(data, mapping = aes(x = "", y = gvh)) + geom_violin(alpha=0.2) + geom_jitter(width = 0.1, height = 0.1, alpha = 0.4, color = "red")
+
+p2 <- ggplot(data, mapping = aes(x = "", y = alm)) + geom_violin(alpha=0.2) + geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "blue")
+
+p3 <- ggplot(data, mapping = aes(x = "", y = mit)) + geom_violin(alpha=0.2) + geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "yellow")
+
+p4 <- ggplot(data, mapping = aes(x = "", y = vac)) + geom_violin(alpha=0.2) + geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "green")
+
+grid.arrange(p1, p2, nrow = 1)
+grid.arrange(p3, p4, nrow = 1)
+```
+
+```{r}
+p1 <- ggplot(data, (mapping = aes(x = loc.site, y = mcg))) + geom_violin(aes(color = loc.site))
+p2 <- ggplot(data, (mapping = aes(x = loc.site, y = gvh))) + geom_violin(aes(color = loc.site))
+
+grid.arrange(p1, p2, nrow = 2)
+```

### 2022-09-22 Resources/LogJournal.Rmd
Added a density plot but without text

@@
 ---
-title: "Log"
-author: "J.A. Busker"
-output: html_document
+title: 'Log Journal Project Theme 9'
+author: 'J.A. Busker (352905 BFV3)'
+output: pdf_document
 ---
@@
 library(cowplot)
+library(ggpubr)
+library(pander)
+library(ggbiplot)
 ```
@@
 
-1.  Sequence Name: Accession number for the SWISS-PROT database.  
-2.  mcg: McGeoch's method for signal sequence recognition.  
-3.  gvh: von Heijne's method for signal sequence recognition.  
-4.  alm: Score of the ALOM membrane spanning region prediction program.  
-5.  mit: Score of discriminant analysis of the amino acid content of the N-terminal region (20 residues long) of mitochondrial and non-mitochondrial proteins.  
-6.  erl: Presence of "HDEL" substring (thought to act as a signal for retention in the endoplasmic reticulum lumen). Binary attribute.  
-7.  pox: Peroxisomal targeting signal in the C-terminus.  
-8.  vac: Score of discriminant analysis of the amino acid content of vacuolar and extracellular proteins.  
-9.  nuc: Score of discriminant analysis of nuclear localization signals of nuclear and non-nuclear proteins.  
-10. Class Distribution: The class is the localization site. Consisting of (abbreviation (full name) the amount):
+* **Sequence Name**: Accession number for the SWISS-PROT database.  
+* **mcg**: McGeoch's method for signal sequence recognition.  
+* **gvh**: von Heijne's method for signal sequence recognition.  
+* **alm**: Score of the ALOM membrane spanning region prediction program.  
+* **mit**: Score of discriminant analysis of the amino acid content of the N-terminal region (20 residues long) of mitochondrial and non-mitochondrial proteins.  
+* **erl**: Presence of "HDEL" substring (thought to act as a signal for retention in the endoplasmic reticulum lumen). Binary attribute.  
+* **pox**: Peroxisomal targeting signal in the C-terminus.  
+* **vac**: Score of discriminant analysis of the amino acid content of vacuolar and extracellular proteins.  
+* **nuc**: Score of discriminant analysis of nuclear localization signals of nuclear and non-nuclear proteins.  
+* **Class Distribution**: The class is the localization site. Consisting of (abbreviation (full name) the amount):
 
+--- ---------------------------------------------- ---
 CYT (cytosolic or cytoskeletal)                    463  
@@
 ERL (endoplasmic reticulum lumen)                    5  
-
-## EDA of the yeast data
+--- ---------------------------------------------- --
 
@@
 ```{r}
-attr_names <- c("seq.name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac","nuc", "loc.site" )
-data_types <- c("str", "float", "float", "float", "float", "float", "bool", "float", "float", "factor")
+attr_names <- c("seq.name", "mcg", "gvh", "alm", "mit", "erl",
+                "pox", "vac","nuc", "loc.site" )
+
+data_types <- c("str", "float", "float", "float", "float", "float",
+                "bool", "float", "float", "factor")
+
 data_labels <- c("Accession number for the SWISS-PROT database",
@@
                  "Score of the ALOM membrane spanning region prediction program",
-                                    "Score of discriminant analysis of the amino acid content of the N-terminal region",
+                 "Score of discriminant analysis of the amino 
+                 acid content of the N-terminal region",
                  "Presence of 'HDEL' substring",
                  "Peroxisomal targeting signal in the C-terminus",
-                                    "Score of discriminant analysis of the amino acid content of vacuolar and extracellular proteins",
-                                    "Score of discriminant analysis of nuclear localization signals of nuclear and non-nuclear proteins",
+                 "Score of discriminant analysis of the amino acid content 
+                 of vacuolar and extracellular proteins",
+                 "Score of discriminant analysis of nuclear 
+                 localization signals of nuclear and non-nuclear proteins",
                  "The class is the localization site")
-codebook <- data.frame(Name=attr_names, Fullname=data_labels, Datatypes=data_types)
 
-codebook %>%
-  kbl(caption = "Codebook of the data") %>%
-  kable_classic(html_font = "Cambria")
+codebook <- data.frame(Name=attr_names, 
+                       Fullname=data_labels, 
+                       Datatypes=data_types)
+
+pander(codebook)
 ```
@@
 ```{r}
-p1 <- ggplot(data, mapping = aes(x = "", y = gvh)) + geom_violin(alpha=0.2) + geom_jitter(width = 0.1, height = 0.1, alpha = 0.4, color = "red")
+p1 <- ggplot(data, mapping = aes(x = "", y = mcg)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.1, height = 0.1, alpha = 0.4, color = "red")
 
-p2 <- ggplot(data, mapping = aes(x = "", y = alm)) + geom_violin(alpha=0.2) + geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "blue")
+p2 <- ggplot(data, mapping = aes(x = "", y = gvh)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "blue")
 
-p3 <- ggplot(data, mapping = aes(x = "", y = mit)) + geom_violin(alpha=0.2) + geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "yellow")
-
-p4 <- ggplot(data, mapping = aes(x = "", y = vac)) + geom_violin(alpha=0.2) + geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "green")
-
-grid.arrange(p1, p2, nrow = 1)
-grid.arrange(p3, p4, nrow = 1)
+ggarrange(p1, p2, nrow = 1)
 ```
@@
 ```{r}
-p1 <- ggplot(data, (mapping = aes(x = loc.site, y = mcg))) + geom_violin(aes(color = loc.site))
-p2 <- ggplot(data, (mapping = aes(x = loc.site, y = gvh))) + geom_violin(aes(color = loc.site))
+p3 <- ggplot(data, mapping = aes(x = "", y = alm)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.1, height = 0.1, alpha = 0.4, color = "red")
 
-grid.arrange(p1, p2, nrow = 2)
+p4 <- ggplot(data, mapping = aes(x = "", y = mit)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "blue")
+
+ggarrange(p3, p4, nrow = 1)
 ```
+
+```{r}
+p5 <- ggplot(data, mapping = aes(x = "", y = pox)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "yellow")
+
+p6 <- ggplot(data, mapping = aes(x = "", y = vac)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "green")
+
+ggarrange(p5, p6, nrow = 1)
+```
+
+```{r}
+p1 <- ggplot(data, aes(x=mcg)) + geom_density(aes(color=loc.site)) + theme(legend.position = "none")
+p2 <- ggplot(data, aes(x=gvh)) + geom_density(aes(color=loc.site)) + theme(legend.position = "none")
+p3 <- ggplot(data, aes(x=alm)) + geom_density(aes(color=loc.site)) + theme(legend.position = "none")
+p4 <- ggplot(data, aes(x=mit)) + geom_density(aes(color=loc.site)) + theme(legend.position = "none")
+p5 <- ggplot(data, aes(x=pox)) + geom_density(aes(color=loc.site)) + theme(legend.position = "none")
+p6 <- ggplot(data, aes(x=vac)) + geom_density(aes(color=loc.site)) + theme(legend.position = "none")
+ggarrange(p1, p2, p3, p4, p5, p6, ncol=3, nrow=2, common.legend=TRUE, legend="right")
+```

### 2022-09-26 Resources/LogJournal.Rmd
Finished the EDA. Added PCA/MDS and much more. Changed gitignore and added the EDA as pdf

@@
 author: 'J.A. Busker (352905 BFV3)'
-output: pdf_document
+output:
+  pdf_document: 
+    pandoc_args: --highlight=breezedark
 ---
@@
 library(ggbiplot)
+library(ggpubr)
+library(reshape2)
+library(ggfortify)
+library(ggalt)
+library(ggridges)
+library(cluster)
 ```
@@
 ---
+
+## EDA of yeast data: Protein Localization
 
@@
 
+### Codebook
+
 Since there is not a codebook. I created my own codebook.
@@
 ```{r}
+# All the attributes name
 attr_names <- c("seq.name", "mcg", "gvh", "alm", "mit", "erl",
@@
 
+# The description of the labels
 data_labels <- c("Accession number for the SWISS-PROT database",
@@
 
+### Load the data
+
 Now we need to load in the data. And change the column names to the attributes abbreviation, since these are non-existent. Let's give all the columns a name. And let's take a quick look at the data.
@@
 ```{r}
+# Read the file in ass a tibble
 data <- as_tibble(read.table("yeast.data", sep = ""))
@@
 
-We can drop the first columns since it is not necessary.
+The yeast data set contains scores per cellular localization sites.
+
+### Clean the data
+
+We can drop the first columns since it is not necessary. Since the sequence names contribute nothing to create a prediction model.
 
 ```{r}
+# Drop the first column
 data <- data[, -1]
@@
 
-We are going to look for any missing values.
+Are there any missing values? Let's take a quick look.
 
 ```{r}
+# Count all the NA values
 sum(is.na(data))
@@
 
-There are not any missing values.
+There are not any missing values. So nothing needs to be changed for this.
 
-Now we need to take a clearer look at the data set using `str()`
+Now we need to take a clearer look at the data set using `str()`.
 
@@
 ```{r}
+# Transform the last column to a factor
 data$loc.site <- as.factor(data$loc.site)
@@
 As you can see the data type of the column loc.site has been succesfully changed to factors.
+
+There are alot of rows. Let's visualize the amount of each classification.
+
+```{r, fig.cap="A waffle chart of the categorical composition"}
+# A grid to fill in 28x53
+df <- expand.grid(y = 1:28, x = 1:53)
+# Sort the table
+categ_table <- sort(table(data$loc.site), decreasing = T)
+df$category <- factor(rep(names(categ_table), categ_table))
+
+ggplot(df, aes(x = x, y = y, fill = category)) + 
+        geom_tile(color = "black", size = 0.5) +
+        scale_x_continuous(expand = c(0, 0)) +
+        scale_y_continuous(expand = c(0, 0), trans = 'reverse') +
+        scale_fill_brewer(palette = "Set3") + 
+        labs(title="Waffle Chart", subtitle="'Class' of localization") + 
+        xlab(NULL) + ylab(NULL)
+```
+
+\newpage
+
+There are alot of CYT and NUC localization. This probably means that CYT has a greater variation in localisation than ERL.
 
@@
 ```{r}
+# Change every 0.5 value to a 0
 data$erl[data$erl == 0.5] <- 0
@@
 ```{r}
+# Change the datatype to a logical
 data$erl <- as.logical(data$erl)
@@
 
+### Univariate Analysis
+
+Let's take a quick look at the data using `summary()`. We can drop column 5 and 9 for this. Column 5, ERL, is a logical datatype and column 9, loc.site, is a string datatype.
+
 ```{r}
-p1 <- ggplot(data, mapping = aes(x = "", y = mcg)) + geom_violin(alpha=0.2) + 
-  geom_jitter(width = 0.1, height = 0.1, alpha = 0.4, color = "red")
-
-p2 <- ggplot(data, mapping = aes(x = "", y = gvh)) + geom_violin(alpha=0.2) + 
-  geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "blue")
-
-ggarrange(p1, p2, nrow = 1)
+summary(data[, -c(5,9)])
 ```
 
-```{r}
+As you can see all the datapoints are between 0 and 1. Se the data already has been transformed with a min-max normalization.
+
+Let's visualise this with ggplot. Using jitterpoints and a violing plot.
+
+```{r, fig.cap="Boxplot comparing basic statistic for all columns", fig.height=12, fig.width=6}
+p1 <- ggplot(data, mapping = aes(x = "", y = mcg)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.2, alpha = 0.25, height = 0, color = "red") + xlab(NULL)
+p2 <- ggplot(data, mapping = aes(x = "", y = gvh)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.2, alpha = 0.25, height = 0, color = "blue") + xlab(NULL)
 p3 <- ggplot(data, mapping = aes(x = "", y = alm)) + geom_violin(alpha=0.2) + 
-  geom_jitter(width = 0.1, height = 0.1, alpha = 0.4, color = "red")
+  geom_jitter(width = 0.2,alpha = 0.25, height = 0, color = "purple") + xlab(NULL)
+p4 <- ggplot(data, mapping = aes(x = "", y = mit)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.2, alpha = 0.25, height = 0, color = "brown") + xlab(NULL)
+p5 <- ggplot(data, mapping = aes(x = "", y = pox)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.2,alpha = 0.25, height = 0, color = "orange") + xlab(NULL)
+p6 <- ggplot(data, mapping = aes(x = "", y = vac)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.2, alpha = 0.25, height = 0, color = "green") + xlab(NULL)
+p7 <- ggplot(data, mapping = aes(x = "", y = nuc)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.2,alpha = 0.25, height = 0, color = "orange") + xlab(NULL)
 
-p4 <- ggplot(data, mapping = aes(x = "", y = mit)) + geom_violin(alpha=0.2) + 
-  geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "blue")
-
-ggarrange(p3, p4, nrow = 1)
+plot <- ggarrange(p1, p2, p3, p4, p5, p6, p7, nrow = 4, ncol = 2)
+annotate_figure(plot, top = text_grob("Boxplots", face = "bold", size = 14))
 ```
 
+\newpage
+
+Mcg and gvh are normally distributed. Alm and mit are skewed but nothing crazy to worry about.
+Vac and nuc are really skewed however, not yet bad enough to worry about it. Pox is looks weird let's take a detailed look at pox.
+
 ```{r}
-p5 <- ggplot(data, mapping = aes(x = "", y = pox)) + geom_violin(alpha=0.2) + 
-  geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "yellow")
-
-p6 <- ggplot(data, mapping = aes(x = "", y = vac)) + geom_violin(alpha=0.2) + 
-  geom_jitter(width = 0.2, height = 0.1, alpha = 0.4, color = "green")
-
-ggarrange(p5, p6, nrow = 1)
+table(data[, "pox"])
 ```
 
+As you can see almost all the numbers are 0. This does not mean that we just discord this column. Maybe it is very significant if the number is not 0.
+
 ```{r}
-p1 <- ggplot(data, aes(x=mcg)) + geom_density(aes(color=loc.site)) + theme(legend.position = "none")
-p2 <- ggplot(data, aes(x=gvh)) + geom_density(aes(color=loc.site)) + theme(legend.position = "none")
-p3 <- ggplot(data, aes(x=alm)) + geom_density(aes(color=loc.site)) + theme(legend.position = "none")
-p4 <- ggplot(data, aes(x=mit)) + geom_density(aes(color=loc.site)) + theme(legend.position = "none")
-p5 <- ggplot(data, aes(x=pox)) + geom_density(aes(color=loc.site)) + theme(legend.position = "none")
-p6 <- ggplot(data, aes(x=vac)) + geom_density(aes(color=loc.site)) + theme(legend.position = "none")
-ggarrange(p1, p2, p3, p4, p5, p6, ncol=3, nrow=2, common.legend=TRUE, legend="right")
+# Subset the data where pox is larger than 0
+subset(data, pox > 0, select=loc.site)
 ```
+
+If the number is non-zero then the probability is very high that the protein is localised in peroxisomal region.
+
+### Bivariate Analysis
+
+With a heatmap, you can easily see where there are correlations between variables. First let's create a correlation matrix
+
+```{r}
+# Create a cor matrix
+cor_matrix <- cor(data[, -c(5,9)])
+cor_matrix <- as_tibble(cor_matrix)
+# Add a column with the varibale names
+(cor_matrix <- cor_matrix %>% mutate(varnames = all_of(attr_names)[-c(1,6,10)]))
+```
+
+The calculated correlation matrix. This needs to be tranformed to a long matrix.
+
+```{r}
+# Create a long matrix
+(cor_matrix_long <- pivot_longer(data = cor_matrix, 
+                                 cols = all_of(attr_names)[-c(1,6,10)], 
+                                 names_to = "variable", values_to = "cor"))
+```
+
+The long calculated correlation matrix.
+
+```{r, fig.cap="A heatmap pairwise correlation of selected numeric variables"}
+
+ggplot(data = cor_matrix_long, aes(x=varnames, y=variable, fill=cor)) + 
+    geom_tile() + 
+    labs(x=NULL, y=NULL, title="Heatmap Correlation") + 
+    scale_fill_gradient(high = "purple", low = "white" )
+```
+
+\newpage
+
+As seen from the heatmap there is some correlation between mcg and gvh. Let's visualise this.
+
+```{r, warning=FALSE, fig.cap="Scatterplot with trendline with the dependent variables"}
+ggplot(data, aes(x=mcg, y=gvh, color=loc.site)) +
+        labs(x="MCG", y="GVH") +
+    geom_jitter(mapping = aes(color=loc.site), 
+                na.rm=T, width=0.2, height=0.2, 
+                alpha=0.5, shape=16, size=0.8) + 
+    ylim(0,1) + labs(title="Scatterplot and trendline") +
+    geom_smooth(formula = y ~ x, method = "loess")
+```
+
+Every variable goes with a slow upward trend except erl this one has a weird curve.
+
+### Class labels
+
+Now we need to look at how the data correlates with the classes. The heigth of the peak doesn’t matter much, shifted peaks do.
+
+```{r,fig.height=6, fig.width=8, fig.cap="Density plots shows class distinction"}
+p1 <- ggplot(data, aes(x=mcg)) + geom_density(aes(color=loc.site))
+p2 <- ggplot(data, aes(x=gvh)) + geom_density(aes(color=loc.site))
+p3 <- ggplot(data, aes(x=alm)) + geom_density(aes(color=loc.site))
+p4 <- ggplot(data, aes(x=mit)) + geom_density(aes(color=loc.site))
+p5 <- ggplot(data, aes(x=pox)) + geom_density(aes(color=loc.site))
+p6 <- ggplot(data, aes(x=vac)) + geom_density(aes(color=loc.site))
+p7 <- ggplot(data, aes(x=nuc)) + geom_density(aes(color=loc.site))
+plot <- ggarrange(p1, p2, p3, p4, p5, p6, p7, ncol=4, nrow=2, 
+          common.legend=TRUE, legend="right")
+annotate_figure(plot, top = text_grob("Density plots", face = "bold", size = 14))
+```
+
+At pox, vac and nuc, you don't see shifted peaks.
+At mcg and gvh you really see shifted peaks this shows a distribution of the different classes.
+There are small shifted peaks at alm en mit, this shows that there is difference but not so much.
+Let's look at mcg and gvh using ridge lines plot.
+
+```{r, warning=FALSE, message=FALSE, fig.cap="Ridge line plot between mcg and gvh"}
+p1 <- ggplot(data, aes(x=mcg, y=loc.site, fill = loc.site)) + 
+  geom_density_ridges2(rel_min_height = 0.005) + theme_minimal() + 
+  coord_cartesian(clip = "off")
+p2 <- ggplot(data, aes(x=gvh, y=loc.site, fill = loc.site)) + 
+  geom_density_ridges2(rel_min_height = 0.005) + theme_minimal() + 
+  coord_cartesian(clip = "off")
+plot <- ggarrange(p1, p2, common.legend = TRUE)
+annotate_figure(plot, top = text_grob("Ridge line plot", face = "bold", size = 14))
+```
+
+\newpage
+
+Now you can really see that the peaks are shifted.
+
+We need to test the data to see if there is significant difference between it. Using a 1-way ANOVA test.
+
+```{r}
+# Perform a one way anova test
+res.aov <- summary(aov(mcg ~ loc.site, data = data))
+res.aov[[1]]$`Pr(>F)`[1]
+```
+
+The P-value is < 0.05. So there is a significant difference.
+
+### Multivariate analysis
+
+One way to see if groups are clustered is to MDS plot the groups and calculated the distance matrix.
+
+```{r}
+# Create a matrix
+matrix <- with(data, rbind(mcg, gvh, alm, mit, pox, vac, nuc))
+(distmat <- dist(matrix))
+```
+
+Here is the distance matrix.
+
+```{r, fig.cap="Classical (Metric) Multidimensional Scaling", fig.width=4, fig.height=2.5}
+autoplot(cmdscale(distmat, eig = TRUE), shape = FALSE, label = TRUE, label.size = 4)
+```
+
+As seen from above mcg and gvh are the clustered group.
+
+A other plot to show is using PCA.
+
+```{r, fig.cap="PCA plot showing the groups"}
+df <- data[-c(5,9)]
+
+pca_res <- prcomp(df, scale. = TRUE)
+
+autoplot(pca_res, data = data, colour = 'loc.site', loadings = TRUE, loadings.label = TRUE)
+```
+
+As earlier shown mcg and gvh are grouped together.

### 2022-09-27 Paper.Rmd
The first commit of the paper

---
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyr)
library(knitr)
library(kableExtra)
library(pander)
library(ggplot2)
library(gridExtra)
```
# Result

To create a great dataset, changes need to be made to the downloaded dataset. The dataset has been gathered of 1484 yeast sequences from SWISS-PROT using the annotations from YPD. In this section, we are going to talk about the results we found from the dataset.

## Codebook

As it is a large dataset, it is wise to look at the codebook first. However, there is no codebook. Nevertheless, there is a file with more information on all the attributes. This is where the information is extracted to make your own codebook.

Below a table with a description of all the attributes abbreviation, explanation and data types.

\begin{center}
Table 1: *The codebook*
\end{center}

```{r}
attr_names <- c("seq.name", "mcg", "gvh", "alm", "mit", "erl",
                "pox", "vac","nuc", "loc.site" )
data_types <- c("str", "float", "float", "float", "float", "float",
                "logical", "float", "float", "factor")
data_labels <- c("Accession number for the SWISS-PROT database",
                 "McGeoch's method for signal sequence recognition",
                 "von Heijne's method for signal sequence recognition",
                 "Score of the ALOM membrane spanning region prediction program",
                 "Score of discriminant analysis of the amino 
                 acid content of the N-terminal region",
                 "Presence of 'HDEL' substring",
                 "Peroxisomal targeting signal in the C-terminus",
                 "Score of discriminant analysis of the amino acid content 
                 of vacuolar and extracellular proteins",
                 "Score of discriminant analysis of nuclear 
                 localization signals of nuclear and non-nuclear proteins",
                 "The class is the localization site")
codebook <- data.frame(Name=attr_names, 
                       Fullname=data_labels, 
                       Datatypes=data_types)
kbl(codebook, booktabs = T, align = "c") %>%
  kable_styling(full_width = T) %>%
  column_spec(2, width = "10cm")
```

The last column is the sequence's localization site. There are 10 different alternatives for this. Here are the 10 in question: 
5  
\begin{center}
Table 2: *Sequence localization sites*
\end{center}

```{r}
abbrv <- c("CYT", "NUC", "MIT", "ME3", "ME2", "ME1", "EXC", "VAC", "POX", "ERL")
loc.fullname <- c("cytosolic or cytoskeletal", "nuclear", "mitochondrial", "membrane protein, no N-terminal signal", "membrane protein, uncleaved signal", "membrane protein, cleaved signal", "extracellular", "vacuolar", "peroxisomal", "endoplasmic reticulum lumen")
amount <- c(463, 429, 244, 163, 51, 44, 37, 30, 20, 5)

pander(data.frame(Abbreviation=abbrv, 
                  Fullname=loc.fullname, 
                  Amount=amount))
```

There are alot of CYT and NUC localization. ERL localization is the least. There are only 5 of these in the dataset.

## Transform data

The data has to be transformed before it can be worked with.

### 2022-10-03 Paper.Rmd
Finished the results section

@@
 library(gridExtra)
+library(ggpmisc)
+library(ggpubr)
+library(ggridges)
+library(ggfortify)
 ```
@@
 
-\begin{center}
-Table 1: *The codebook*
-\end{center}
-
 ```{r}
@@
                 "pox", "vac","nuc", "loc.site" )
-data_types <- c("str", "float", "float", "float", "float", "float",
-                "logical", "float", "float", "factor")
+data_types <- c("str", "double", "double", "double", "double", "double",
+                "logical", "double", "double", "factor")
 data_labels <- c("Accession number for the SWISS-PROT database",
@@
 
-The last column is the sequence's localization site. There are 10 different alternatives for this. Here are the 10 in question: 
-5  
 \begin{center}
-Table 2: *Sequence localization sites*
+\emph{Table 1:} The codebook.
 \end{center}
+
+The last column is the sequence's localization site.
+There are 10 different possibilities for this.
+Yeast protein were classified into 10 classes: **cytoplasmic:** cytoskeletal (CYT); nuc]ear (NUC); vacuolar (VAC); mitochondrial (MIT); isomal (POX); **extracellular:** including those localized against the cell wall (EXC); proteins localized to the lumen of the endoplasmic reticulum (ERL); membrane proteins with a cleaved signal (ME1); membrane proteins with an uncleared signal (ME2); and membrane proteins with no N-terminal sign (ME3), where ME1, ME2,and ME3 proteins may be localized to the plasma membrane, the endoplasmic reticulum membrane, or the membrane of a golgi body.
+
+Here are the 10 in question:
 
@@
 amount <- c(463, 429, 244, 163, 51, 44, 37, 30, 20, 5)
-
 pander(data.frame(Abbreviation=abbrv, 
@@
 
+\begin{center}
+\emph{Table 2:} Sequence localization sites.
+\end{center}
+
 There are alot of CYT and NUC localization. ERL localization is the least. There are only 5 of these in the dataset.
 
-## Transform data
+## Dataset
 
-The data has to be transformed before it can be worked with.
+8 features were used in classification: the presence or absence of an HDEL pattern as a signal for retention in the endoplasmic reticulum lumen (erl); the result of discriminant analysis on the amino acid content of vacuolar and extracellular proteins (vac); the result discriminant analysis on the amino acid composition of the 20-residue N-terminal region of mitochondrial and non-mitochondrial proteins (mit); the presence absence of nuclear localization consensus patterns combined with a term reflecting the frequency of basic residues (nuc); and some combination of the presence of a short sequence motif and the result of discriminant analysis of the amino acid composition of the protein sequence (pox).
+
+```{r}
+data <- as_tibble(read.table("yeast.data", sep = ""))
+colnames(data) <- attr_names
+pander(head(data))
+```
+
+\begin{center}
+\emph{Table 3:} First 6 rows of the dataset.
+\end{center}
+
+As seen from the table the loaded dataset. Has a char datatype for the first and last column. And every other column has a double as datatype.
+
+There are `r nrow(data)` rows and `r ncol(data)` columns.
+
+\newpage
+
+```{r, fig.cap="A waffle chart of the categorical composition", fig.width=6, fig.height=3}
+df <- expand.grid(y = 1:28, x = 1:53)
+categ_table <- sort(table(data$loc.site), decreasing = T)
+df$category <- factor(rep(names(categ_table), categ_table))
+ggplot(df, aes(x = x, y = y, fill = category)) +
+geom_tile(color = "black", size = 0.5) +
+scale_x_continuous(expand = c(0, 0)) +
+scale_y_continuous(expand = c(0, 0), trans = 'reverse') +
+scale_fill_brewer(palette = "Set3") +
+labs(title="Waffle Chart", subtitle="'Class' of localization") +
+xlab(NULL) + ylab(NULL)
+```
+
+There are alot of CYT and NUC localization. And there are only `r nrow(subset(data, loc.site=="ERL", loc.site))` of ERL.
+
+## Clean the data
+
+To ensure the quality of the data. The data had to be transformed before it could be worked with.
+
+The first column has been dropped since it is not necessary. Since the sequence names contribute nothing to create a prediction model.
+
+```{r}
+data <- data[, -1]
+pander(head(data))
+```
+
+\begin{center}
+\emph{Table 4:} First 6 rows of the dataset with the first column dropped.
+\end{center}
+
+As stated on the website there are `r sum(is.na(data))` missing values.
+
+```{r}
+data$loc.site <- as.factor(data$loc.site)
+str(data$loc.site)
+```
+
+The data type of the column loc.site has been succesfully changed to factors.
+
+The ERL column has changed. From an double to a logical datatype. As seen in figure 2 below, only numbers 1 and 0.5 appear, but for a logical datatype you need 1 and 0. All 0.5 have been changed to 0.
+
+\newpage
+
+```{r, warning=FALSE, message=FALSE, fig.cap="ERL column data distrubution", fig.width=10, fig.height=4}
+my_list <- as.data.frame(table(data[,"erl"]))
+names(my_list) <- c("Values", "Freq")
+ggplot(data, mapping = aes(x = "", y = erl)) + geom_violin(alpha=0.2) + geom_jitter(width = 0.5, alpha = 0.25, height = 0.005, color = "red") + xlab(NULL) + annotate(geom = "table", x = 0.5, y = 0.75, label = list(my_list)) + ylim(0, 1) + labs(title="ERL Variable", subtitle="Data distrubution")
+```
+
+```{r}
+data$erl[data$erl == 0.5] <- 0
+data$erl <- as.logical(data$erl)
+m <- sapply(data, typeof)
+m[m=="integer"] <- "factor"
+pander(m)
+```
+
+\begin{center}
+\emph{Table 5:} Each column name and the typeof the datatype.
+\end{center}
+
+Each column has been changed so that it has the right data type.
+
+## Exploratory data analysis
+
+```{r}
+table <- summary(data[, -c(5,9)])
+panderOptions('table.continues', '')
+pander(table)
+```
+
+\begin{center}
+\emph{Table 6:} Summary of the dataset.
+\end{center}
+
+As observed in table 6 the data has already been transformed with a min-max normalisation.
+All the datapoints are between 0 and 1.Mcg and gvh are normally distributed.
+Alm and mit are skewed but nothing to crazy to be worried about.
+Vac and nuc are really skewed however, not yet bad enough to worry about it.
+Pox is very oddly distributed.
+
+\newpage
+
+```{r, fig.cap="Closer look at the pox variable", fig.width=10, fig.height=4}
+my_list <- as.data.frame(table(data[,"pox"]))
+names(my_list) <- c("Values", "Freq")
+p1 <- ggplot(data, mapping = aes(x = "", y = pox)) + geom_violin(alpha=0.2) + 
+  geom_jitter(width = 0.5, alpha = 0.25, height = 0.005, color = "red") + xlab(NULL)
+p1 + annotate(geom = "table", x = 0.5, y = 0.75, label = list(my_list)) + labs(title="Pox Variable", subtitle="Data distrubution")
+```
+
+Almost all the datapoint are 0. This could cause problems later on. Since more than 95% of all the datapoints are 0.
+
+```{r, fig.cap="Frequency plot of pox variablee", fig.height=3}
+d <- as.data.frame(table(subset(data, pox > 0, select=loc.site)))
+ggplot(d, aes(Var1, Freq)) +
+  geom_linerange(
+    aes(x = Var1, ymin = 0, ymax = Freq), 
+    color = "lightgray", size = 1.5
+    )+
+  geom_point(aes(color = Var1), size = 2) + labs(title="Frequency plot of pox variable", x="localization site", y = "Count")
+```
+
+If the pox value is not 0 then the classification is Pox, Mit or Cyt. This means that Pox has a high probability when the Pox value is not 0.
+
+With a heatmap, you can easily see where there are correlations between variables.
+
+```{r}
+cor_matrix <- cor(data[, -c(5,9)])
+cor_matrix <- as_tibble(cor_matrix)
+cor_matrix <- cor_matrix %>% mutate(varnames = all_of(attr_names)[-c(1,6,10)])
+kbl(cor_matrix, booktabs = T, align = "c") %>%
+  kable_styling(full_width = T)
+```
+
+\begin{center}
+\emph{Table 7:} The calculated correlation matrix.
+\end{center}
+
+This has been tranformed to a long matrix.
+
+```{r}
+cor_matrix_long <- pivot_longer(data = cor_matrix, 
+                                 cols = all_of(attr_names)[-c(1,6,10)], 
+                                 names_to = "variable", values_to = "cor")
+pander(head(cor_matrix_long))
+```
+
+\begin{center}
+\emph{Table 8:} The long calculated correlation matrix.
+\end{center}
+
+```{r, fig.cap="A heatmap pairwise correlation of selected numeric variables", fig.height=4}
+
+ggplot(data = cor_matrix_long, aes(x=varnames, y=variable, fill=cor)) + 
+    geom_tile() + 
+    labs(x=NULL, y=NULL, title="Heatmap Correlation") + 
+    scale_fill_gradient(high = "purple", low = "white" )
+```
+
+As seen from the heatmap there is some correlation between mcg and gvh. While on the contrary Pox barely has any correlation with a other variable.
+
+There needs to be data correlation with the classes. The heigth of the peak doesn’t matter much, shifted peaks do.
+
+```{r, warning=FALSE, message=FALSE, fig.cap="Ridge line plot between mcg and gvh"}
+p1 <- ggplot(data, aes(x=mcg, y=loc.site, fill = loc.site)) + 
+  geom_density_ridges2(rel_min_height = 0.005) + theme_minimal() + 
+  coord_cartesian(clip = "off")
+p2 <- ggplot(data, aes(x=gvh, y=loc.site, fill = loc.site)) + 
+  geom_density_ridges2(rel_min_height = 0.005) + theme_minimal() + 
+  coord_cartesian(clip = "off")
+plot <- ggarrange(p1, p2, common.legend = TRUE)
+annotate_figure(plot, top = text_grob("Ridge line plot", face = "bold", size = 14))
+```
+
+There are shifted peaks. this shows a distribution of the different classes.
+
+A 1-way ANOVA test was carried out on the data.
+
+```{r}
+# Perform a one way anova test
+res.aov <- summary(aov(mcg ~ loc.site, data = data))
+p.value <- res.aov[[1]]$`Pr(>F)`[1]
+```
+
+The P-value: `r p.value` is < 0.05. So there is a significant difference.
+
+\newpage
+
+```{r, fig.cap="Classical (Metric) Multidimensional Scaling", fig.width=4, fig.height=2.5}
+matrix <- with(data, rbind(mcg, gvh, alm, mit, pox, vac, nuc))
+distmat <- dist(matrix)
+autoplot(cmdscale(distmat, eig = TRUE), shape = FALSE, label = TRUE, label.size = 4)
+```
+
+As found in the heatmap Mcg and Gvh are clustered. Vac and Alm are somewhat clustered they are close to eachother. Nuc, Mit and Pox are all on their own.
+
+```{r, fig.cap="PCA plot", fig.width=6, fig.height=3}
+df <- data[-c(5,9)]
+
+pca_res <- prcomp(df, scale. = TRUE)
+
+autoplot(pca_res, data = data, colour = 'loc.site', loadings = TRUE, loadings.label = TRUE)
+```
+
+Mcg and Gvh are also clustered in this PCA plot. As in contrast with the MDS plot Vac and Alm are not clustered at all. Pox is in the centre.
+
+\newpage
+
+# Discussion
+
+# Conclusion

### 2022-10-03 Resources/LogJournal.Rmd
Finished the results section

@@
 library(ggbiplot)
-library(ggpubr)
 library(reshape2)

### 2022-10-04 Paper.Rmd
Finished the discussion and conclusion. Added the research question to the readme

@@
 
-There are alot of CYT and NUC localization. And there are only `r nrow(subset(data, loc.site=="ERL", loc.site))` of ERL.
+Figure 1 shows the amount of each clasification. There are alot of CYT and NUC localization. And there are only `r nrow(subset(data, loc.site=="ERL", loc.site))` of ERL.
 
@@
 
-# Discussion
+# Discussion & Conclusion
 
-# Conclusion
+Our research question is: Can machine learning predict protein localisation based on their aminoacid sequences.
+And to answer this question properly, a clean dataset is needed to create a model.
+The results will show what is wrong with the data and the correlations between the different classifications.
+The results are not completely trustworthy because there is too little of them.
+Namely, there are only 1484 oberservations.
+This is too few for prediction with 10 different classifications.
+
+## Discussion
+
+Nor was the Erl column considered.
+This indicates the presence of "HDEL" substring.
+If this had been looked at, more correlations would have emerged.
+Which could later become significant.
+There was also something wrong with the ERL column.
+As stated on the site, it is a binary/logical type.
+Either 0 or 1.
+It is present or it is not present.
+However, there were only values of 0.5 and 1.
+All values of 0.5 were changed to 0.
+
+In the case of the Pox column, something strange was noticed.
+Out of 1484 datapoints, 15 are not zero, 11 of them have as classifications Pox.
+This means that if the value is not zero that the classification is highly probable pox.
+
+## Conclusion
+
+The aim was to create a clean dataset, to be used for our machine learning project.
+There is a clean dataset with no missing values and the correct data types.
+The data gives a clear overview of a protein classification and their attributes.
+However, not much was available.
+With only 1484 observations, this is just too little.
+And there is a very big difference between the number of observations per classifications.
+For ERL there are only 5 observations and for CYT there are 463.
+There is too big a difference between these.
+By improving this dataset over time, the prediction can be perfected even more.
+These constant improvements are important to find.

### 2022-10-04 Paper.Rmd
Fixed some grammer/spelling errors

@@
 To create a great dataset, changes need to be made to the downloaded dataset.
-The dataset has been gathered of 1484 yeast sequences from SWISS-PROT using the annotations from YPD.
+The dataset consists of 1484 yeast sequences from SWISS-PROT using the annotations from YPD.
 In this section, we are going to talk about the results we found from the dataset.
@@
 
-Below a table with a description of all the attributes abbreviation, explanation and data types.
+Below is a table with a description of all the attributes' abbreviations, explanations, and data types.
 
@@
 The last column is the sequence's localization site.
-There are 10 different possibilities for this.
-Yeast protein were classified into 10 classes: **cytoplasmic:** cytoskeletal (CYT); nuc]ear (NUC); vacuolar (VAC); mitochondrial (MIT); isomal (POX); **extracellular:** including those localized against the cell wall (EXC); proteins localized to the lumen of the endoplasmic reticulum (ERL); membrane proteins with a cleaved signal (ME1); membrane proteins with an uncleared signal (ME2); and membrane proteins with no N-terminal sign (ME3), where ME1, ME2,and ME3 proteins may be localized to the plasma membrane, the endoplasmic reticulum membrane, or the membrane of a golgi body.
+There are ten different possibilities for this.
+Yeast proteins were classified into ten classes: **cytoplasmic:** cytoskeletal (CYT); nuc]ear (NUC); vacuolar (VAC); mitochondrial (MIT); isomal (POX); **extracellular:** including those localized against the cell wall (EXC); proteins localized to the lumen of the endoplasmic reticulum (ERL); membrane proteins with a cleaved signal (ME1); membrane proteins with an uncleared signal (ME2); and membrane proteins with no N-terminal sign (ME3), where ME1, ME2,and ME3 proteins may be localized to the plasma membrane, the endoplasmic reticulum membrane, or the membrane of a golgi body.
 
-Here are the 10 in question:
+Here are the ten in question:
 
@@
 
-There are alot of CYT and NUC localization. ERL localization is the least. There are only 5 of these in the dataset.
+There are alot of CYT and NUC localizations. ERL localization is the least. There are only five of these in the dataset.
 
@@
 
-8 features were used in classification: the presence or absence of an HDEL pattern as a signal for retention in the endoplasmic reticulum lumen (erl); the result of discriminant analysis on the amino acid content of vacuolar and extracellular proteins (vac); the result discriminant analysis on the amino acid composition of the 20-residue N-terminal region of mitochondrial and non-mitochondrial proteins (mit); the presence absence of nuclear localization consensus patterns combined with a term reflecting the frequency of basic residues (nuc); and some combination of the presence of a short sequence motif and the result of discriminant analysis of the amino acid composition of the protein sequence (pox).
+Eight features were used in classification: the presence or absence of an HDEL pattern as a signal for retention in the endoplasmic reticulum lumen (erl); The results of discriminant analysis on the amino acid content of vacuolar and extracellular proteins (vac); the result of discriminant analysis on the amino acid composition of the 20-residue N-terminal region of mitochondrial and non-mitochondrial proteins (mit); the presence or absence of nuclear localization consensus patterns combined with a term reflecting the frequency of basic residues (nuc); and some combination of the presence of a short sequence motif and the result of discriminant analysis of the amino acid composition of the protein sequence (pox).
 
@@
 \begin{center}
-\emph{Table 3:} First 6 rows of the dataset.
+\emph{Table 3:} First six rows of the dataset.
 \end{center}
 
-As seen from the table the loaded dataset. Has a char datatype for the first and last column. And every other column has a double as datatype.
+As seen from the table, the loaded dataset. It has a char datatype for the first and last column. And every other column has a double datatype.
 
@@
 
-Figure 1 shows the amount of each clasification. There are alot of CYT and NUC localization. And there are only `r nrow(subset(data, loc.site=="ERL", loc.site))` of ERL.
+Figure 1 shows the number of each classification of the dataset. There are alot of CYT and NUC localization. And there are only `r nrow(subset(data, loc.site=="ERL", loc.site))` of ERL.
 
@@
 
-The first column has been dropped since it is not necessary. Since the sequence names contribute nothing to create a prediction model.
+The first column has been dropped since it is not necessary. Since the sequence names contribute nothing to creating a prediction model.
 
@@
 \begin{center}
-\emph{Table 4:} First 6 rows of the dataset with the first column dropped.
+\emph{Table 4:} First six rows of the dataset with the first column dropped.
 \end{center}
@@
 
-The data type of the column loc.site has been succesfully changed to factors.
+The data type of the column loc.site has been successfully changed to factors.
 
-The ERL column has changed. From an double to a logical datatype. As seen in figure 2 below, only numbers 1 and 0.5 appear, but for a logical datatype you need 1 and 0. All 0.5 have been changed to 0.
+The ERL column has changed. from a double to a logical datatype. As seen in figure 2 below, only numbers 1 and 0.5 appear, but for a logical datatype you need 1 and 0. All 0.5 have been changed to 0.
 
@@
 As observed in table 6 the data has already been transformed with a min-max normalisation.
-All the datapoints are between 0 and 1.Mcg and gvh are normally distributed.
-Alm and mit are skewed but nothing to crazy to be worried about.
-Vac and nuc are really skewed however, not yet bad enough to worry about it.
+All the datapoints are between 0 and 1.
+Mcg and gvh are normally distributed.
+Alm and mit are skewed, but nothing too crazy to be worried about.
+Vac and nuc are really skewed.
+However, it's not yet bad enough to worry about it.
 Pox is very oddly distributed.
@@
 
-Almost all the datapoint are 0. This could cause problems later on. Since more than 95% of all the datapoints are 0.
+Almost all the data points are 0. This could cause problems later on. Since more than 95% of all the data points are 0.
 
@@
 
-If the pox value is not 0 then the classification is Pox, Mit or Cyt. This means that Pox has a high probability when the Pox value is not 0.
+If the pox value is not zero then the classification is Pox, Mit or Cyt. This means that Pox has a high probability when the Pox value is not zero.
 
@@
 
-This has been tranformed to a long matrix.
+This has been transformed into a long matrix.
 
@@
 
-As seen from the heatmap there is some correlation between mcg and gvh. While on the contrary Pox barely has any correlation with a other variable.
+As seen from the heatmap, there is some correlation between mcg and gvh. Pox, on the other hand, has almost no correlation with any other variable.
 
-There needs to be data correlation with the classes. The heigth of the peak doesn’t matter much, shifted peaks do.
+There needs to be data correlation with the classes. The height of the peak doesn’t matter much, shifted peaks do.
 
@@
 
-There are shifted peaks. this shows a distribution of the different classes.
+There are shifted peaks. This shows the distribution of the different classes.
 
@@
 
-As found in the heatmap Mcg and Gvh are clustered. Vac and Alm are somewhat clustered they are close to eachother. Nuc, Mit and Pox are all on their own.
+As found in the heatmap, Mcg and Gvh are clustered. Vac and Alm are somewhat clustered; they are close to each other. Nuc, Mit and Pox are all on their own.
 
@@
 
-Mcg and Gvh are also clustered in this PCA plot. As in contrast with the MDS plot Vac and Alm are not clustered at all. Pox is in the centre.
+Mcg and Gvh are also clustered in this PCA plot. As such, in contrast with the MDS plot, Vac and Alm are not clustered at all. Pox is in the centre.
 
@@
 
-Our research question is: Can machine learning predict protein localisation based on their aminoacid sequences.
+The research question is: can machine learning predict protein localization based on their aminoacid sequences?
 And to answer this question properly, a clean dataset is needed to create a model.
@@
 Nor was the Erl column considered.
-This indicates the presence of "HDEL" substring.
+This indicates the presence of the "HDEL" substring.
 If this had been looked at, more correlations would have emerged.
@@
 In the case of the Pox column, something strange was noticed.
-Out of 1484 datapoints, 15 are not zero, 11 of them have as classifications Pox.
-This means that if the value is not zero that the classification is highly probable pox.
+Out of 1484 datapoints, 15 are not zero.
+11 of them have the classification of Pox.
+This means that if the value is not zero, the classification is highly probable Pox.
 
@@
 
-The aim was to create a clean dataset, to be used for our machine learning project.
+The aim was to create a clean dataset to be used for our machine learning project.
 There is a clean dataset with no missing values and the correct data types.
-The data gives a clear overview of a protein classification and their attributes.
+The data gives a clear overview of the protein classification and their attributes.
 However, not much was available.
 With only 1484 observations, this is just too little.
-And there is a very big difference between the number of observations per classifications.
-For ERL there are only 5 observations and for CYT there are 463.
+And there is a very big difference between the number of observations per classification.
+For ERL there are only 5 observations, and for CYT there are 463.
 There is too big a difference between these.

### 2022-10-10 Resources/LogJournal.Rmd
Chosen which MIL ALGO in labjournal

@@
 As earlier shown mcg and gvh are grouped together.
+
+\newpage
+
+# MIL algorithms
+
+First we need to investigate which standard ML algorithms is the best to use for this dataset using Weka.
+
+Weka accepts .arff files. You can convert a .csv to a .arff file in Weka. So we need to export the clean dataset to a .csv file.
+
+```{r, eval=FALSE}
+write.table(data,"data.csv", quote = FALSE, sep = ",", row.names=FALSE)
+```
+
+Open the .csv in Weka under tools -> ArffViewer. And saved the file as .arff. The first 15 lines of the .arff file needs to look like this.
+
+```{r}
+cat(readLines("data.arff", n = 15), sep = "\n")
+```
+
+*Criteria for MIL algorithms*
+
+An accurate algorithm is needed that accurately predicts the dataset.
+It needs a high percentage of correct classifications on the dataset.
+An algorithm that is fast.
+What if there is a dataset of 1 million observations.
+And it should also not contain too many false negatives and false positives.
+
+Research has been done on which algorithm is the best.
+From every representatives of all classifier categories.
+And ZeroR and OneR for a baseline performance.
+Eg: Decision Trees (C4.5, J48, RandomForest), Nearest Neighbor (IBk), SVM (SMO), NaiveBayes (Naïve Bayes) and Linear Logistic (SimpleLogistic).
+And carried out classifications 10-fold cross validation.
+
+```{r}
+wekaResult <- read.table("ResultWeka.txt", header = TRUE)
+colnames(wekaResult) <- c("Classifier", "TP Rate Avg.", "FP Rate Avg.", 
+                          "Precision", "VAC Precision")
+pander(wekaResult)
+```
+
+Several algorithms cannot make a prediction for the classification VAC (0.000) and for some they get an error (?).
+1 algorithm IBk has a very low prediction for VAC.
+All the algorithm above in table NUMBER except for ZeroR, OneR, SMO and SimpleLogistic were used in the Weka Experimenter to compare.
+
+![Ranking of chosen algorithms]("images/RankingAlgo.png")
+
+This shows the ranking of the algorithms by the number of times a given algorithm beat the other algorithms. And RandomForest beat all 4 algorithms.
+
+![Ranking of chosen algorithms by SD]("images/WekaSD.png")
+
+RandomForest has a 'v' next to their result. This means that the difference in the accuracy for these algorithms compared to all the other algorithms is significant.
+
+After testing multiple algorithms in Weka, the algorithm RandomForest was chosen.
+It has the highest average precision of 0.611.
+And this was one of the few algorithms where the precision for VAC was not a '?'.
+
+But for the ratings, no algorithm could make a prediction except 1 for VAC, however, it was also very low. So we looked at whether there is siginificant difference if we remove this classifier.
+
+![Removed class difference]("images/WekaDataEdit.png")
+
+And by the looks of it, there is no significant difference. So VAC is not removed from the dataset.

### 2022-10-13 Resources/LogJournal.Rmd
Finishes the logjournal

@@
 And by the looks of it, there is no significant difference. So VAC is not removed from the dataset.
+
+RandomForest has seven attributes. Below is a table with each attribute and its meaning.
+
+```{r, echo=FALSE}
+attr <- c("P", "I", "num-slots", "K", "M", "V", "S")
+attr.descr <- c("Size of each bag, as a percentage of the training set size.", 
+                "Number of iterations.", "Number of execution slots.", 
+                "Number of attributes to randomly investigate.", 
+                "Number of attributes to randomly investigate.", 
+                "Set minimum numeric class variance proportion.", 
+                "Seed for random number generator.")
+pander(data.frame(Attribute=attr, Description=attr.descr))
+```
+
+Weka will find the optimal parameters for a classifier. With CVParameterSelection in Weka were the best parameters found.
+
+```{r, echo=FALSE}
+old.val <- c(100, 100, 1, 0, 1, 0.001, 1)
+new.val <- c(50, 1000, 0, 1, 1, 0.001, 1)
+pander(data.frame(Attribute=attr, Old.Values=old.val, New.Values=new.val))
+```
+
+This table shows the parameters with their default settings and their new.
+P, I, num-slots and K were different than the default settings.
+The build will take more time because it does 10 times more I(terations).
+
+![Iterations ranked]("images/WekaIterRank.png")
+
+Number 1 in the figure above is with 1000 I(terations) and number 2 is 100 I(terations). And there is not a significant difference. So for performance, the value for I will be 100.
+
+Weka Meta learners were used to get a better idea of an optimal classifier.
+
+```{r, echo=FALSE}
+tbl <- read.table("MetaResults.txt", header = TRUE)
+colnames(tbl) <- c("Classifier", "TP Rate", "FP Rate", "Precision", "ROC Area", " ")
+pander(tbl)
+```
+
+As it can be seen in this table, RandomForest with tweaked parameters is the best. This one has no ? at VAC and precision.
+
+Since there are 10 classifications, ROC curve have been caught from one classifications.
+
+\newpage
+
+![ROC Curve (NUC)]("images/ROCCurve.png")
+
+This is a nice curve with a relative low ROC Curve. This means that this classification can correspond well to a particular decision threshold.

### 2022-10-14 Resources/LogJournal.Rmd
Added barplot in the MIL Section (Did not knit)

@@
 
+The experimenter can save the results in a csv file.
+
 ```{r}
-wekaResult <- read.table("ResultWeka.txt", header = TRUE)
-colnames(wekaResult) <- c("Classifier", "TP Rate Avg.", "FP Rate Avg.", 
-                          "Precision", "VAC Precision")
-pander(wekaResult)
+weka.result <- read.table("Result.csv", header = TRUE, sep = ",")
+# Remove weka.classifiers.* from classifier name
+for (i in 1:length(weka.result$Key_Scheme)) {
+  name <- weka.result$Key_Scheme[i]
+  tweaked.name <- strsplit(name, split = ".", fixed = TRUE)[[1]][4]
+  weka.result$Key_Scheme[i] <- tweaked.name
+}
 ```
+
+Now the results of the csv can been plotted with ease of use. Like the ROC and precision.
+
+```{r, fig.cap="Area under the curve for different classifiers"}
+# Create a dataframe of the mean of the ROC of every classifier
+res <- aggregate(Area_under_ROC ~ Key_Scheme, data = weka.result, mean)
+
+ggplot(data=res, aes(x=Key_Scheme, y=Area_under_ROC, fill=Key_Scheme)) + 
+  geom_bar(stat="identity") + 
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
+  labs(x="Classifier", y="Area under curve (ROC)", 
+       title="Area under the curve for different classifiers") + 
+  geom_hline(aes(yintercept=min(Area_under_ROC)), linetype = 'dashed') + 
+  geom_hline(aes(yintercept=max(Area_under_ROC)), linetype = 'dashed') + 
+  geom_text(aes(label=round(Area_under_ROC, digits=2)), vjust=-0.3, size=3.5)
+```
+
+RandomForest has the highest area under the curve, while OneR has the lowest with 0.49 and ZeroR with a close second last with 0.5.
+
+```{r, fig.cap="Average precisionfor different classifiers"}
+res <- aggregate(IR_precision ~ Key_Scheme, data = weka.result, mean)
+
+ggplot(data=res, aes(x=Key_Scheme, y=IR_precision, fill=Key_Scheme)) + 
+  geom_bar(stat="identity") + 
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
+  labs(x="Classifier", y="Precision", 
+       title="Average precisionfor different classifiers") + 
+  geom_hline(aes(yintercept=min(IR_precision)), linetype = 'dashed') + 
+  geom_hline(aes(yintercept=max(IR_precision)), linetype = 'dashed') + 
+  geom_text(aes(label=round(IR_precision, digits=2)), vjust=-0.3, size=3.5)
+```
+
+RandomForest also has the highest average precision with 0.66. And OneR has the lowest with 0.04. ZeroR is not in this plots because it only consist of NaN.
 
@@
 1 algorithm IBk has a very low prediction for VAC.
-All the algorithm above in table NUMBER except for ZeroR, OneR, SMO and SimpleLogistic were used in the Weka Experimenter to compare.
+All the algorithm above in the figure except for ZeroR, OneR, SMO and SimpleLogistic were used in the Weka Experimenter to compare.
 
@@
 As it can be seen in this table, RandomForest with tweaked parameters is the best. This one has no ? at VAC and precision.
-
-Since there are 10 classifications, ROC curve have been caught from one classifications.
-
-\newpage
-
-![ROC Curve (NUC)]("images/ROCCurve.png")
-
-This is a nice curve with a relative low ROC Curve. This means that this classification can correspond well to a particular decision threshold.

### 2022-10-24 Resources/LogJournal.Rmd
Finished the logjournal

@@
 
-```{r}
+```{r, warning=FALSE}
 # Create a cor matrix
@@
 # Add a column with the varibale names
-(cor_matrix <- cor_matrix %>% mutate(varnames = all_of(attr_names)[-c(1,6,10)]))
+cor_matrix <- cor_matrix %>% mutate(varnames = all_of(attr_names)[-c(1,6,10)])
+kbl(cor_matrix, booktabs = T, align = "c") %>%
+  kable_styling(full_width = T)
 ```
@@
 
-```{r}
+```{r, warning=FALSE}
 # Create a long matrix
-(cor_matrix_long <- pivot_longer(data = cor_matrix, 
-                                 cols = all_of(attr_names)[-c(1,6,10)], 
-                                 names_to = "variable", values_to = "cor"))
+cols <- all_of(attr_names)[-c(1,6,10)]
+cor_matrix_long <- pivot_longer(data = cor_matrix, 
+                                 cols = cols,
+                                 names_to = "variable", values_to = "cor")
+pander(head(cor_matrix_long))
 ```
@@
 
+\newpage
+
 ```{r, fig.cap="PCA plot showing the groups"}
@@
 
-*Criteria for MIL algorithms*
+**Criteria for MIL algorithms**
 
@@
   geom_bar(stat="identity") + 
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
   labs(x="Classifier", y="Area under curve (ROC)", 
@@
   geom_hline(aes(yintercept=max(Area_under_ROC)), linetype = 'dashed') + 
-  geom_text(aes(label=round(Area_under_ROC, digits=2)), vjust=-0.3, size=3.5)
+  geom_text(aes(label=round(Area_under_ROC, digits=2)), vjust=1.5, size=3.5)
 ```
@@
 RandomForest has the highest area under the curve, while OneR has the lowest with 0.49 and ZeroR with a close second last with 0.5.
+
+\newpage
 
@@
   geom_bar(stat="identity") + 
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
   labs(x="Classifier", y="Precision", 
-       title="Average precisionfor different classifiers") + 
+       title="Average precision for different classifiers") + 
   geom_hline(aes(yintercept=min(IR_precision)), linetype = 'dashed') + 
   geom_hline(aes(yintercept=max(IR_precision)), linetype = 'dashed') + 
-  geom_text(aes(label=round(IR_precision, digits=2)), vjust=-0.3, size=3.5)
+  geom_text(aes(label=round(IR_precision, digits=2)), vjust=1.5, size=3.5)
 ```
@@
 
-As it can be seen in this table, RandomForest with tweaked parameters is the best. This one has no ? at VAC and precision.
+As it can be seen in this table, RandomForest with tweaked parameters is the best. This one has no '?' at VAC and precision.

### 2022-10-24 LogJournal.Rmd
Moved the logjournal file

--- 2022-10-24 Resources/LogJournal.Rmd
@@
 # Read the file in ass a tibble
-data <- as_tibble(read.table("yeast.data", sep = ""))
+data <- as_tibble(read.table("Resources/yeast.data", sep = ""))
 colnames(data) <- attr_names
@@
 ```{r}
-cat(readLines("data.arff", n = 15), sep = "\n")
+cat(readLines("Resources/data.arff", n = 15), sep = "\n")
 ```
@@
 ```{r}
-weka.result <- read.table("Result.csv", header = TRUE, sep = ",")
+weka.result <- read.table("Resources/Result.csv", header = TRUE, sep = ",")
 # Remove weka.classifiers.* from classifier name
@@
 
-![Ranking of chosen algorithms]("images/RankingAlgo.png")
+![Ranking of chosen algorithms]("Resources/images/RankingAlgo.png")
 
@@
 
-![Ranking of chosen algorithms by SD]("images/WekaSD.png")
+![Ranking of chosen algorithms by SD]("Resources/images/WekaSD.png")
 
@@
 
-![Removed class difference]("images/WekaDataEdit.png")
+![Removed class difference]("Resources/images/WekaDataEdit.png")
 
@@
 
-![Iterations ranked]("images/WekaIterRank.png")
+![Iterations ranked]("Resources/images/WekaIterRank.png")
 
@@
 ```{r, echo=FALSE}
-tbl <- read.table("MetaResults.txt", header = TRUE)
+tbl <- read.table("Resources/MetaResults.txt", header = TRUE)
 colnames(tbl) <- c("Classifier", "TP Rate", "FP Rate", "Precision", "ROC Area", " ")

### 2022-10-28 LogJournal.Rmd
Added a TOC and finished the logjournal

@@
 ---
-title: 'Log Journal Project Theme 9'
-author: 'J.A. Busker (352905 BFV3)'
 output:
@@
 
-```{r setup, include=FALSE}
+```{r, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE, comment=NA)
+```
+
+<!-- Copyright (c) 2022 John Busker. -->
+<!-- Licensed under GPLv3. See gpl.md -->
+
+# Machine Learning Research log
+
+## Can machine learning predict protein localization based on their aminoacid sequences
+
+\vspace*{\fill}
+\begin{flushright}
+John Busker\\
+352905\\
+Bio-Informatics\\
+Dave Langers, and Bart Barnard\\
+28-10-2022
+\end{flushright}
+
+\newpage
+
+# Table of Contents
+
+1. [R Environment](#r-environment) \dotfill 3
+
+2. [EDA](#eda) \dotfill 4  
+  2.1 [About The Data](#about-the-data) \dotfill 4  
+  2.2 [Codebook](#codebook) \dotfill 5  
+  2.3 [Loading The Data](#loading-the-data) \dotfill 6  
+  2.4 [Cleaning The Data](#cleaning-the-data) \dotfill 6  
+  2.5 [Data Transformation](#data-transformation) \dotfill 9  
+  2.6 [Univariate Analysis](#univariate-analysis) \dotfill 9  
+  2.7 [Bivariate Analysis](#bivariate-analysis) \dotfill 12   
+  2.8 [Class Labels](#class-labels) \dotfill 15  
+  2.9 [Multivariate Analysis](#multivariate-analysis) \dotfill 18
+
+3. [MIL Algorithms](#mil-algorithms) \dotfill 20  
+  3.1 [Criteria For MIL Algorithms](#criteria-for-mil-algorithms) \dotfill 20
+
+\newpage
+
+# R Environment
+
+Multiple libraries have been uses for data analysis and further processes. The libraries used are:
+
+```{r setup, message=FALSE}
 library(tidyr)
@@
 
-<!-- Copyright (c) 2022 John Busker. -->
-<!-- Licensed under GPLv3. See gpl.md -->
+\newpage
 
----
+# EDA
 
-## EDA of yeast data: Protein Localization
+## About The Data
 
-The data has been retrived from [uci.](https://archive.ics.uci.edu/ml/datasets/Yeast) The dataset has been gathered of 1484 yeast sequences from SWISS-PROT using the annotations from YPD.
+The data has been retrieved from [uci.](https://archive.ics.uci.edu/ml/datasets/Yeast) The dataset consists of 1484 yeast sequences from SWISS-PROT using the annotations from YPD.
+In this section, we are going to talk about the results we found from the dataset.
+Eight features were used in classification: the presence or absence of an HDEL pattern as a signal for retention in the endoplasmic reticulum lumen (erl); The results of discriminant analysis on the amino acid content of vacuolar and extracellular proteins (vac); the result of discriminant analysis on the amino acid composition of the 20-residue N-terminal region of mitochondrial and non-mitochondrial proteins (mit); the presence or absence of nuclear localization consensus patterns combined with a term reflecting the frequency of basic residues (nuc); and some combination of the presence of a short sequence motif and the result of discriminant analysis of the amino acid composition of the protein sequence (pox).
 
@@
 
-### Codebook
+Yeast proteins were classified into ten classes: **cytoplasmic:** cytoskeletal (CYT); nuc]ear (NUC); vacuolar (VAC); mitochondrial (MIT); isomal (POX); **extracellular:** including those localized against the cell wall (EXC); proteins localized to the lumen of the endoplasmic reticulum (ERL); membrane proteins with a cleaved signal (ME1); membrane proteins with an uncleared signal (ME2); and membrane proteins with no N-terminal sign (ME3), where ME1, ME2,and ME3 proteins may be localized to the plasma membrane, the endoplasmic reticulum membrane, or the membrane of a golgi body.
 
-Since there is not a codebook. I created my own codebook.
+## Codebook
+
+Since there is not a codebook. A own codebook has been created
 
@@
 
-### Load the data
+## Loading The Data
 
@@
 
-The yeast data set contains scores per cellular localization sites.
+The yeast data set contains scores per cellular localization sites. The last column is the sequence's localization site. There are ten different possibilities for this.
 
-### Clean the data
+## Cleaning The Data
 
@@
 
+## Data Transformation
+
 The ERL variable also should be changed. As you can see, it is now a num data type but I needs to be a bool/binary datatype. Let's take a closer look at th ERL column.
@@
 
-### Univariate Analysis
+## Univariate Analysis
 
@@
 
-### Bivariate Analysis
+## Bivariate Analysis
 
@@
 
-The long calculated correlation matrix.
+The long calculated correlation matrix. Let's visualise this using a heatmap.
 
@@
 
-### Class labels
+## Class Labels
 
@@
 
-### Multivariate analysis
+## Multivariate analysis
 
@@
 
-# MIL algorithms
+# MIL Algorithms
 
@@
 
-**Criteria for MIL algorithms**
+## Criteria For MIL Algorithms
 

### 2022-11-02 Paper.Rmd
FInished the paper

@@
 ---
-output: pdf_document
+output:
+  pdf_document: 
+    pandoc_args: --highlight=breezedark
 ---
+
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = TRUE, comment=NA)
 library(tidyr)
-library(knitr)
 library(kableExtra)
-library(pander)
 library(ggplot2)
 library(gridExtra)
-library(ggpmisc)
+library(cowplot)
 library(ggpubr)
+library(pander)
+library(ggbiplot)
+library(reshape2)
+library(ggfortify)
+library(ggalt)
 library(ggridges)
-library(ggfortify)
+library(cluster)
 ```
-# Result
+
+<!-- Copyright (c) 2022 John Busker. -->
+
+<!-- Licensed under GPLv3. See gpl.md -->
+
+# Cellular Localization Prediction
+
+## Can the cellular localization Sites of proteins be predicted using probabilistic classification.
+
+![<https://www.hj9l.com/zhidao/changshi/13854.html>](Resources/images/cover.jpg)
+
+```{=tex}
+\vspace*{\fill}
+\begin{flushright}
+John Busker\\
+352905\\
+Bio-Informatics\\
+Dave Langers, and Bart Barnard\\
+28-10-2022
+\end{flushright}
+```
+\newpage
+
+# Table of Contents
+
+1. [Introduction](#introduction) \dotfill 3
+
+2. [Materials and Methods](#materials-and-methods) \dotfill 4  
+3. [Results](#results) \dotfill 5  
+  3.1 [Codebook](#codebook) \dotfill 5  
+  3.2 [Machine Learning](#machine-learning) \dotfill 8
+
+4. [Discussion and Conclusion](#discussion-and-conclusion) \dotfill 12  
+  4.1 [Discussion](#discussion) \dotfill 12  
+  4.2 [Conclusion](#conclusion) \dotfill 12
+
+5. [References](#references) \dotfill 13
+
+\newpage
+
+# Introduction
+
+Protein activities and their subcellular locations in cells are strongly connected.
+Finding proteins' positions within cells is frequently the initial step in understanding them, and their sites serve as a model for medication design.
+The amino acid sequence of a protein mostly determines where it will localize within a cell.
+Using this knowledge, a software has been developed to define a simple model of classification of yeast localization in cells.
+The efficient arrangement of biological data into databases and the creation of algorithms, such string algorithms, to manipulate that data, have largely defined the area of computational biology.
+The dataset consists of 1484 yeast sequences from SWISS-PROT using the annotations from YPD.
+
+Research question: Can the cellular localization Sites of proteins be predicted using probabilistic classification.
+
+\newpage
+
+# Materials and Methods
+
+The data has been retrieved from [uci.](https://archive.ics.uci.edu/ml/datasets/Yeast) The dataset consists of 1484 yeast sequences from SWISS-PROT using the annotations from YPD.
+In this section, we are going to talk about the results we found from the dataset.
+
+Yeast proteins were classified into ten classes: **cytoplasmic:** cytoskeletal (CYT); nuc]ear (NUC); vacuolar (VAC); mitochondrial (MIT); isomal (POX); **extracellular:** including those localized against the cell wall (EXC); proteins localized to the lumen of the endoplasmic reticulum (ERL); membrane proteins with a cleaved signal (ME1); membrane proteins with an uncleared signal (ME2); and membrane proteins with no N-terminal sign (ME3), where ME1, ME2,and ME3 proteins may be localized to the plasma membrane, the endoplasmic reticulum membrane, or the membrane of a golgi body.
+
+Eight features were used in classification. Here below are all the features
+
+-   **mcg**: McGeoch's method for signal sequence recognition.\
+-   **gvh**: von Heijne's method for signal sequence recognition.\
+-   **alm**: Score of the ALOM membrane spanning region prediction program.\
+-   **mit**: Score of discriminant analysis of the amino acid content of the N-terminal region (20 residues long) of mitochondrial and non-mitochondrial proteins.\
+-   **erl**: Presence of "HDEL" substring (thought to act as a signal for retention in the endoplasmic reticulum lumen). Binary attribute.\
+-   **pox**: Peroxisomal targeting signal in the C-terminus.\
+-   **vac**: Score of discriminant analysis of the amino acid content of vacuolar and extracellular proteins.\
+-   **nuc**: Score of discriminant analysis of nuclear localization signals of nuclear and non-nuclear proteins.
+
+It is always important to look at the data and to look at the quality of the data.
+No proper conclusion can be drawn from low quality data.
+For recurrent errors the data can be corrected.
+The data needs to be modified to make sure the data gets the highest quality to perform the analysis on.
+This is achieved by performing exploratory data analysis on the dataset.
+Modifying and removing parts of the dataset to get best.
+A RandomForest classifier has been chosen for the classification of the unknown instances.
+The parameters has been optimized for usability and speed.
+See the [logjournal](https://github.com/AlfonsoJan/Theme9).
+
+A java application has been developed.
+The software using Weka carried out the calculations required to carry out the probabilistic inference.
+The application can quickly classify new instances given from a file on from the command line.
+See the [repo](https://bitbucket.org/janalfonso/wekawrapper/) for the code.
+
+\newpage
+
+# Results
 
@@
 
-```{r}
+```{r, echo=FALSE}
 attr_names <- c("seq.name", "mcg", "gvh", "alm", "mit", "erl",
@@
 
+```{=tex}
 \begin{center}
@@
 \end{center}
-
+```
 The last column is the sequence's localization site.
 There are ten different possibilities for this.
-Yeast proteins were classified into ten classes: **cytoplasmic:** cytoskeletal (CYT); nuc]ear (NUC); vacuolar (VAC); mitochondrial (MIT); isomal (POX); **extracellular:** including those localized against the cell wall (EXC); proteins localized to the lumen of the endoplasmic reticulum (ERL); membrane proteins with a cleaved signal (ME1); membrane proteins with an uncleared signal (ME2); and membrane proteins with no N-terminal sign (ME3), where ME1, ME2,and ME3 proteins may be localized to the plasma membrane, the endoplasmic reticulum membrane, or the membrane of a golgi body.
+Yeast proteins were classified into ten classes: **cytoplasmic:** CYT; NUC; VAC; MIT; POX; **extracellular:** EXC; ERL; ME1; ME2; ME3.
 
@@
 
-```{r}
+```{r, echo=FALSE}
 abbrv <- c("CYT", "NUC", "MIT", "ME3", "ME2", "ME1", "EXC", "VAC", "POX", "ERL")
@@
 \end{center}
-
+```
 There are alot of CYT and NUC localizations. ERL localization is the least. There are only five of these in the dataset.
@@
 
-```{r}
-data <- as_tibble(read.table("yeast.data", sep = ""))
+```{r, echo=FALSE}
+data <- as_tibble(read.table("Resources/yeast.data", sep = ""))
 colnames(data) <- attr_names
@@
 \end{center}
-
+```
 As seen from the table, the loaded dataset. It has a char datatype for the first and last column. And every other column has a double datatype.
@@
 
-\newpage
-
-```{r, fig.cap="A waffle chart of the categorical composition", fig.width=6, fig.height=3}
+```{r, fig.cap="A waffle chart of the categorical composition", fig.width=6, fig.height=3, echo=FALSE}
 df <- expand.grid(y = 1:28, x = 1:53)
@@
 
-Figure 1 shows the number of each classification of the dataset. There are alot of CYT and NUC localization. And there are only `r nrow(subset(data, loc.site=="ERL", loc.site))` of ERL.
-
-## Clean the data
+Figure 2 shows the number of each classification of the dataset. There are alot of CYT and NUC localization. And there are only `r nrow(subset(data, loc.site=="ERL", loc.site))` of ERL.
 
@@
 
-```{r}
+```{r, echo=FALSE}
 data <- data[, -1]
@@
 \end{center}
+```
+As stated on the website there are `r sum(is.na(data))` missing values. This has been confirmed using R.
 
-As stated on the website there are `r sum(is.na(data))` missing values.
-
-```{r}
+```{r, echo=FALSE}
 data$loc.site <- as.factor(data$loc.site)
-str(data$loc.site)
 ```
 
-The data type of the column loc.site has been successfully changed to factors.
+It can be more convenient to have a variable in a factor form instead of numeric form. The data type of the column loc.site has been successfully changed to factors.
 
-The ERL column has changed. from a double to a logical datatype. As seen in figure 2 below, only numbers 1 and 0.5 appear, but for a logical datatype you need 1 and 0. All 0.5 have been changed to 0.
+The ERL column has changed. from a double to a logical datatype. As seen in table 5 below, only numbers 1 and 0.5 appear, but for a logical datatype you need 1 and 0. All 0.5 have been changed to 0.
 
-\newpage
-
-```{r, warning=FALSE, message=FALSE, fig.cap="ERL column data distrubution", fig.width=10, fig.height=4}
-my_list <- as.data.frame(table(data[,"erl"]))
-names(my_list) <- c("Values", "Freq")
-ggplot(data, mapping = aes(x = "", y = erl)) + geom_violin(alpha=0.2) + geom_jitter(width = 0.5, alpha = 0.25, height = 0.005, color = "red") + xlab(NULL) + annotate(geom = "table", x = 0.5, y = 0.75, label = list(my_list)) + ylim(0, 1) + labs(title="ERL Variable", subtitle="Data distrubution")
+```{r, echo=FALSE}
+pander(table(data$erl))
 ```
 
-```{r}
+```{=tex}
+\begin{center}
+\emph{Table 5:} A table showing distrubtion of the ERL column.
+\end{center}
+```
+```{r, echo=FALSE}
 data$erl[data$erl == 0.5] <- 0
@@
 
+```{=tex}
 \begin{center}
-\emph{Table 5:} Each column name and the typeof the datatype.
+\emph{Table 6:} Each column name and the typeof the datatype.
 \end{center}
-
+```
 Each column has been changed so that it has the right data type.
 
-## Exploratory data analysis
-
-```{r}
-table <- summary(data[, -c(5,9)])
-panderOptions('table.continues', '')
-pander(table)
-```
-
-\begin{center}
-\emph{Table 6:} Summary of the dataset.
-\end{center}
-
-As observed in table 6 the data has already been transformed with a min-max normalisation.
-All the datapoints are between 0 and 1.
-Mcg and gvh are normally distributed.
-Alm and mit are skewed, but nothing too crazy to be worried about.
-Vac and nuc are really skewed.
-However, it's not yet bad enough to worry about it.
-Pox is very oddly distributed.
+At pox, vac and nuc, you don't see shifted peaks.
+At mcg and gvh you really see shifted peaks this shows a distribution of the different classes.
+There are small shifted peaks at alm en mit, this shows that there is difference but not so much.
 
@@
 
-```{r, fig.cap="Closer look at the pox variable", fig.width=10, fig.height=4}
-my_list <- as.data.frame(table(data[,"pox"]))
-names(my_list) <- c("Values", "Freq")
-p1 <- ggplot(data, mapping = aes(x = "", y = pox)) + geom_violin(alpha=0.2) + 
-  geom_jitter(width = 0.5, alpha = 0.25, height = 0.005, color = "red") + xlab(NULL)
-p1 + annotate(geom = "table", x = 0.5, y = 0.75, label = list(my_list)) + labs(title="Pox Variable", subtitle="Data distrubution")
-```
-
-Almost all the data points are 0. This could cause problems later on. Since more than 95% of all the data points are 0.
-
-```{r, fig.cap="Frequency plot of pox variablee", fig.height=3}
-d <- as.data.frame(table(subset(data, pox > 0, select=loc.site)))
-ggplot(d, aes(Var1, Freq)) +
-  geom_linerange(
-    aes(x = Var1, ymin = 0, ymax = Freq), 
-    color = "lightgray", size = 1.5
-    )+
-  geom_point(aes(color = Var1), size = 2) + labs(title="Frequency plot of pox variable", x="localization site", y = "Count")
-```
-
-If the pox value is not zero then the classification is Pox, Mit or Cyt. This means that Pox has a high probability when the Pox value is not zero.
-
-With a heatmap, you can easily see where there are correlations between variables.
-
-```{r}
-cor_matrix <- cor(data[, -c(5,9)])
-cor_matrix <- as_tibble(cor_matrix)
-cor_matrix <- cor_matrix %>% mutate(varnames = all_of(attr_names)[-c(1,6,10)])
-kbl(cor_matrix, booktabs = T, align = "c") %>%
-  kable_styling(full_width = T)
-```
-
-\begin{center}
-\emph{Table 7:} The calculated correlation matrix.
-\end{center}
-
-This has been transformed into a long matrix.
-
-```{r}
-cor_matrix_long <- pivot_longer(data = cor_matrix, 
-                                 cols = all_of(attr_names)[-c(1,6,10)], 
-                                 names_to = "variable", values_to = "cor")
-pander(head(cor_matrix_long))
-```
-
-\begin{center}
-\emph{Table 8:} The long calculated correlation matrix.
-\end{center}
-
-```{r, fig.cap="A heatmap pairwise correlation of selected numeric variables", fig.height=4}
-
-ggplot(data = cor_matrix_long, aes(x=varnames, y=variable, fill=cor)) + 
-    geom_tile() + 
-    labs(x=NULL, y=NULL, title="Heatmap Correlation") + 
-    scale_fill_gradient(high = "purple", low = "white" )
-```
-
-As seen from the heatmap, there is some correlation between mcg and gvh. Pox, on the other hand, has almost no correlation with any other variable.
-
-There needs to be data correlation with the classes. The height of the peak doesn’t matter much, shifted peaks do.
-
-```{r, warning=FALSE, message=FALSE, fig.cap="Ridge line plot between mcg and gvh"}
+```{r, warning=FALSE, message=FALSE, fig.cap="Ridge line plot between mcg and gvh", echo=FALSE, fig.width=6, fig.height=6}
 p1 <- ggplot(data, aes(x=mcg, y=loc.site, fill = loc.site)) + 
@@
 
-There are shifted peaks. This shows the distribution of the different classes.
+Now you can really see that the peaks are shifted.
 
-A 1-way ANOVA test was carried out on the data.
+We need to test the data to see if there is significant difference between it. Using a 1-way ANOVA test.
 
-```{r}
+```{r, echo=FALSE}
 # Perform a one way anova test
@@
 
-The P-value: `r p.value` is < 0.05. So there is a significant difference.
+The P-value is `r p.value` \< 0.05. So there is a significant difference.
+
+## Machine Learning
+
+Weka was used to test which algorithm is the best to use for the classification of instances.
+An accurate algorithm is needed that accurately predicts the dataset.
+It needs a high percentage of correct classifications on the dataset.
+An algorithm needs to be fast for the best usage.
+What if there is a dataset of 1 million observations.
+And it should also not contain too many false negatives and false positives.
+
+Using the experiment of Weka testing has been done on every representatives of all classifier categories.
+And ZeroR and OneR for a baseline performance.
+Decision Trees (C4.5, J48, RandomForest), Nearest Neighbor (IBk), SVM (SMO), NaiveBayes (Naïve Bayes) and Linear Logistic (SimpleLogistic).
+And carried out classifications 10-fold cross validation.
+The result were saved in a .csv file.
+
+```{r, echo=FALSE}
+weka.result <- read.table("Resources/Result.csv", header = TRUE, sep = ",")
+# Remove weka.classifiers.* from classifier name
+for (i in 1:length(weka.result$Key_Scheme)) {
+  name <- weka.result$Key_Scheme[i]
+  tweaked.name <- strsplit(name, split = ".", fixed = TRUE)[[1]][4]
+  weka.result$Key_Scheme[i] <- tweaked.name
+}
+```
+
+```{r,fig.cap="Area under the curve for different classifiers", echo=FALSE}
+res <- aggregate(Area_under_ROC ~ Key_Scheme, data = weka.result, mean)
+ggplot(data=res, aes(x=Key_Scheme, y=Area_under_ROC, fill=Key_Scheme)) + 
+  geom_bar(stat="identity") + 
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
+  labs(x="Classifier", y="Area under curve (ROC)", 
+       title="Area under the curve for different classifiers") + 
+  geom_hline(aes(yintercept=min(Area_under_ROC)), linetype = 'dashed') + 
+  geom_hline(aes(yintercept=max(Area_under_ROC)), linetype = 'dashed') + 
+  geom_text(aes(label=round(Area_under_ROC, digits=2)), vjust=1.5, size=3.5)
+```
+
+An integrated measurement of a measurably effect or phenomena is the area under the curve.
+A higher Area under the curve the better.
+RandomForest has the highest area under the curve, while OneR has the lowest with 0.49 and ZeroR with a close second last with 0.5.
+
+The algorithm also needs te be precise.
 
@@
 
-```{r, fig.cap="Classical (Metric) Multidimensional Scaling", fig.width=4, fig.height=2.5}
-matrix <- with(data, rbind(mcg, gvh, alm, mit, pox, vac, nuc))
-distmat <- dist(matrix)
-autoplot(cmdscale(distmat, eig = TRUE), shape = FALSE, label = TRUE, label.size = 4)
+```{r, fig.cap="Average precisionfor different classifiers", echo=FALSE}
+res <- aggregate(IR_precision ~ Key_Scheme, data = weka.result, mean)
+
+ggplot(data=res, aes(x=Key_Scheme, y=IR_precision, fill=Key_Scheme)) + 
+  geom_bar(stat="identity") + 
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
+  labs(x="Classifier", y="Precision", 
+       title="Average precision for different classifiers") + 
+  geom_hline(aes(yintercept=min(IR_precision)), linetype = 'dashed') + 
+  geom_hline(aes(yintercept=max(IR_precision)), linetype = 'dashed') + 
+  geom_text(aes(label=round(IR_precision, digits=2)), vjust=1.5, size=3.5)
+
 ```
 
-As found in the heatmap, Mcg and Gvh are clustered. Vac and Alm are somewhat clustered; they are close to each other. Nuc, Mit and Pox are all on their own.
+RandomForest also has the highest average precision with 0.66. And OneR has the lowest with 0.04. ZeroR is not in this plots because it only consist of NaN.
 
-```{r, fig.cap="PCA plot", fig.width=6, fig.height=3}
-df <- data[-c(5,9)]
+Several algorithms cannot make a prediction for the classification VAC (0.000) and for some they get an error (?).
+1 algorithm IBk has a very low prediction for VAC.
+All the algorithm above in the figure except for ZeroR, OneR, SMO and SimpleLogistic were used in the Weka Experimenter to compare.
 
-pca_res <- prcomp(df, scale. = TRUE)
+![Ranking of chosen algorithms](%22Resources/images/WekaSD.png%22)
 
-autoplot(pca_res, data = data, colour = 'loc.site', loadings = TRUE, loadings.label = TRUE)
-```
+RandomForest has a 'v' next to their result.
+This means that the difference in the accuracy for these algorithms compared to all the other algorithms is significant.
+And RandomForest beat all 4 algorithms.
 
-Mcg and Gvh are also clustered in this PCA plot. As such, in contrast with the MDS plot, Vac and Alm are not clustered at all. Pox is in the centre.
+The classification of VAC seems to be the most concerning of all. It has been removed and tested against not removed. To see if there is an significant difference.
 
@@
 
-# Discussion & Conclusion
+![Removed class difference](%22Resources/images/WekaDataEdit.png%22)
+
+There is no significant difference between removing VAC and keeping VAC.
+
+Weka will find the optimal parameters for a classifier. With CVParameterSelection in Weka were the best parameters found.
+
+```{r, echo=FALSE}
+attr <- c("P", "I", "num-slots", "K", "M", "V", "S")
+attr.descr <- c("Size of each bag, as a percentage of the training set size.", 
+                "Number of iterations.", "Number of execution slots.", 
+                "Number of attributes to randomly investigate.", 
+                "Number of attributes to randomly investigate.", 
+                "Set minimum numeric class variance proportion.", 
+                "Seed for random number generator.")
+old.val <- c(100, 100, 1, 0, 1, 0.001, 1)
+new.val <- c(50, 1000, 0, 1, 1, 0.001, 1)
+pander(data.frame(Attribute=attr, Description=attr.descr, Old.Values=old.val, New.Values=new.val))
+```
+
+```{=tex}
+\begin{center}
+\emph{Table 7:} Each parameter with the old and new best value if necessary.
+\end{center}
+```
+This table shows the parameters with their default settings and their new.
+P, I, num-slots and K were different than the default settings.
+The build will take more time because it does 10 times more I(terations).
+
+![Iterations ranked]("Resources/images/WekaIterRank.png")
+
+Number 1 in the figure above is with 1000 I(terations) and number 2 is 100 I(terations). And there is not a significant difference.
+
+\newpage
+
+# Discussion and Conclusion
 
@@
 
+Only 8 classifiers had been used.
+From our research, we found that RandomForest is the best, however, perhaps another untested classifier is much better than RandomForest.
+Only the removal of VAC was considered what if much better precision and classification would have run if another classifier had been removed.
+
 ## Conclusion
@@
 These constant improvements are important to find.
+
+1484 yeast proteins were classified into 10 classes with an accuracy of 61%.
+
+\newpage
+
+# References

### 2022-11-09 Paper.Rmd
Added the project proposal

@@
 
-5. [References](#references) \dotfill 13
+5. [Project Proposal](#project-proposal) \dotfill 13
+
+6. [References](#references) \dotfill 14
 
@@
 Protein activities and their subcellular locations in cells are strongly connected.
-Finding proteins' positions within cells is frequently the initial step in understanding them, and their sites serve as a model for medication design.
+Finding proteins’ positions within cells is frequently the initial step in understanding them, and their sites serve as a model for medication design.
 The amino acid sequence of a protein mostly determines where it will localize within a cell.
-Using this knowledge, a software has been developed to define a simple model of classification of yeast localization in cells.
-The efficient arrangement of biological data into databases and the creation of algorithms, such string algorithms, to manipulate that data, have largely defined the area of computational biology.
+Using this knowledge, a software program has been developed to define a simple model for the classification of yeast localization in cells.
+The efficient arrangement of biological data into databases and the creation of algorithms, such as string algorithms, to manipulate that data have largely defined the area of computational biology.
 The dataset consists of 1484 yeast sequences from SWISS-PROT using the annotations from YPD.
 
-Research question: Can the cellular localization Sites of proteins be predicted using probabilistic classification.
+Research question: How accurate can the cellular localization sites of proteins be predicted using probabilistic classification?
 
@@
 It is always important to look at the data and to look at the quality of the data.
-No proper conclusion can be drawn from low quality data.
-For recurrent errors the data can be corrected.
-The data needs to be modified to make sure the data gets the highest quality to perform the analysis on.
+No proper conclusion can be drawn from low-quality data.
+For recurrent errors, the data can be corrected.
+The data needs to be modified to make sure it is of the highest quality to perform the analysis on.
 This is achieved by performing exploratory data analysis on the dataset.
-Modifying and removing parts of the dataset to get best.
+Modifying and removing parts of the dataset to get the best possible dataset.
 A RandomForest classifier has been chosen for the classification of the unknown instances.
-The parameters has been optimized for usability and speed.
-See the [logjournal](https://github.com/AlfonsoJan/Theme9).
+The parameters have been optimized for usability and speed.
+See the [logjournal](https://github.com/AlfonsoJan/Theme9) [1].
 
-A java application has been developed.
+A Java application has been developed.
 The software using Weka carried out the calculations required to carry out the probabilistic inference.
-The application can quickly classify new instances given from a file on from the command line.
-See the [repo](https://bitbucket.org/janalfonso/wekawrapper/) for the code.
+The application can quickly classify new instances given from a file or from the command line.
+See the [repo](https://bitbucket.org/janalfonso/wekawrapper/) [2] for the code.
 
@@
 The dataset consists of 1484 yeast sequences from SWISS-PROT using the annotations from YPD.
-In this section, we are going to talk about the results we found from the dataset.
+In this section, we are going to talk about the results we found from the dataset [3].
 
@@
 
-Eight features were used in classification: the presence or absence of an HDEL pattern as a signal for retention in the endoplasmic reticulum lumen (erl); The results of discriminant analysis on the amino acid content of vacuolar and extracellular proteins (vac); the result of discriminant analysis on the amino acid composition of the 20-residue N-terminal region of mitochondrial and non-mitochondrial proteins (mit); the presence or absence of nuclear localization consensus patterns combined with a term reflecting the frequency of basic residues (nuc); and some combination of the presence of a short sequence motif and the result of discriminant analysis of the amino acid composition of the protein sequence (pox).
+Eight features were used in classification: the presence or absence of an HDEL pattern as a signal for retention in the endoplasmic reticulum lumen (erl); The results of discriminant analysis on the amino acid content of vacuolar and extracellular proteins (vac); the result of discriminant analysis on the amino acid composition of the 20-residue N-terminal region of mitochondrial and non-mitochondrial proteins (mit); the presence or absence of nuclear localization consensus patterns combined with a term reflecting the frequency of basic residues (nuc); and some combination of the presence of a short sequence motif and the result of discriminant analysis of the amino acid composition of the protein sequence (pox) [3].
 
@@
 
-Weka was used to test which algorithm is the best to use for the classification of instances.
+Weka was used to test which algorithm is best to use for the classification of instances.
 An accurate algorithm is needed that accurately predicts the dataset.
@@
 An algorithm needs to be fast for the best usage.
-What if there is a dataset of 1 million observations.
-And it should also not contain too many false negatives and false positives.
+What if there is a dataset of 1 million observations?
+And it should also not contain too many false negatives or false positives.
 
@@
 
-An integrated measurement of a measurably effect or phenomena is the area under the curve.
-A higher Area under the curve the better.
-RandomForest has the highest area under the curve, while OneR has the lowest with 0.49 and ZeroR with a close second last with 0.5.
+The area under the curve is a way to measure an effect or phenomenon as a whole.
+The greater the area under the curve, the better.
+RandomForest has the largest area under the curve, while OneR has the lowest with 0.49 and ZeroR is a close second with 0.5.
 
@@
 
-RandomForest also has the highest average precision with 0.66. And OneR has the lowest with 0.04. ZeroR is not in this plots because it only consist of NaN.
+RandomForest also has the highest average precision, at 0.66. OneR has the lowest, at 0.04. ZeroR is not in this plot because it only consists of NaN.
 
-Several algorithms cannot make a prediction for the classification VAC (0.000) and for some they get an error (?).
-1 algorithm IBk has a very low prediction for VAC.
-All the algorithm above in the figure except for ZeroR, OneR, SMO and SimpleLogistic were used in the Weka Experimenter to compare.
+Several algorithms cannot make a prediction for the classification VAC (0.000), and for some they get an error (?).
+The IBk algorithm has a very low prediction for VAC.
+Except for ZeroR, OneR, SMO, and SimpleLogistic, all of the algorithms in the figure above were compared in the Weka experimenter.
 
@@
 
-The research question is: can machine learning predict protein localization based on their aminoacid sequences?
+The research question is: can machine learning predict protein localization based on their amino acid sequences?
 And to answer this question properly, a clean dataset is needed to create a model.
 The results will show what is wrong with the data and the correlations between the different classifications.
-The results are not completely trustworthy because there is too little of them.
+The results are not completely trustworthy because there are too few of them.
 Namely, there are only 1484 oberservations.
-This is too few for prediction with 10 different classifications.
+With ten different classifications, this is insufficient for prediction.
 
@@
 In the case of the Pox column, something strange was noticed.
-Out of 1484 datapoints, 15 are not zero.
+Out of 1484 data points, 15 are not zero.
 11 of them have the classification of Pox.
@@
 
-Only 8 classifiers had been used.
+Only eight classifiers had been used.
 From our research, we found that RandomForest is the best, however, perhaps another untested classifier is much better than RandomForest.
-Only the removal of VAC was considered what if much better precision and classification would have run if another classifier had been removed.
+Only the removal of VAC was considered, even though much better precision and classification would have been achieved if another classifier had been removed.
 
@@
 There is a clean dataset with no missing values and the correct data types.
-The data gives a clear overview of the protein classification and their attributes.
+The data provides a clear picture of protein classification and attributes.
 However, not much was available.
@@
 
+# Project Proposal
+
+I will be taking the minor in Application Design.
+And we have been using the program Weka this quarter, and when you open the program, you immediately see how clustered and old-fashioned this program is.
+My project proposal is to design a modern Weka interface.
+
+\newpage
+
 # References
+
+[1] = J.A. Busker, Theme 9, (2022), GitHub repository, https://github.com/AlfonsoJan/Theme9
+
+[2] = J.A. Busker, Weka Wrapper, (2022), BitBucket repository, https://bitbucket.org/janalfonso/wekawrapper
+
+[3] = P. Horton, K. Nakai, A Probabilistic Classification System for Predicting the Cellular Localization Sites of Proteins, (1996), AAAI, https://www.aaai.org/Papers/ISMB/1996/ISMB96-012.pdf
